---
title: "Journal (reproducible report)"
author: "Yasser Makki"
date: "2020-11-24"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

# Chapter 2: Tidyverse

Last compiled: `r Sys.Date()`

## Load libraries

```{r}
library(tidyverse)
library(lubridate)
library(readxl)
library("writexl")
```


## Importing Files
```{r}
bikes_tbl      <- read_excel(path = "00_data/01_bike_sales/01_raw_data/bikes.xlsx")
orderlines_tbl <- read_excel("00_data/01_bike_sales/01_raw_data/orderlines.xlsx")
bikeshops_tbl  <- read_excel("00_data/01_bike_sales/01_raw_data/bikeshops.xlsx")

```


## Examining Data
```{r}
orderlines_tbl
glimpse(orderlines_tbl)

```


## Joining Data
```{r}
left_join(orderlines_tbl, bikes_tbl, by = c("product.id" = "bike.id"))

bike_orderlines_joined_tbl <- orderlines_tbl %>%
  left_join(bikes_tbl, by = c("product.id" = "bike.id")) %>%
  left_join(bikeshops_tbl, by = c("customer.id" = "bikeshop.id"))
bike_orderlines_joined_tbl %>% glimpse()
```


## Wrangling Data
```{r}
bike_orderlines_joined_tbl %>% 
  select(category) %>%
  filter(str_detect(category, "^Mountain")) %>% 
  unique()

bike_orderlines_wrangled_city_separated_tbl <- bike_orderlines_joined_tbl %>%
  
  separate(col    = category,
           into   = c("category.1", "category.2", "category.3"),
           sep    = " - ") %>% 
  
  separate(col    = location,
           into   = c("City", "State"),
           sep    = ", ") %>% 
  
  mutate(total.price = price * quantity)%>%
  
  select(-...1, -gender)%>%
  
  select(-ends_with(".id"))%>%
  
  bind_cols(bike_orderlines_joined_tbl %>% select(order.id)) %>%
  
  select(order.id, contains("order"), contains("model"), contains("category"),
         price, quantity, total.price,
         everything()) %>%
  
  rename(bikeshop = name) %>%
  
  set_names(names(.) %>% 
              
              str_replace_all("\\.", "_"))

```



## Business Insights
### Sales by location

#### Manipulate

```{r}
sales_by_location_tbl <- bike_orderlines_wrangled_city_separated_tbl %>%
  
  select(State, City, total_price) %>%
  
  group_by(State) %>%
  
  summarize(state_sales = sum(total_price)) %>%
  
  mutate(sales_formatted = scales::dollar(state_sales, big.mark = ".", 
                                          decimal.mark = ",", 
                                          prefix = "", 
                                          suffix = " €"))
```



### Sales by location and year

#### Manipulate

```{r}
sales_by_state_year_tbl <- bike_orderlines_wrangled_city_separated_tbl %>%
  
  select(order_date, total_price, State) %>%
  mutate(year = year(order_date)) %>%
  
  group_by(State, year) %>%
  summarise(sales = sum(total_price)) %>%
  ungroup() %>%
  
  mutate(sales_formatted = scales::dollar(sales, big.mark = ".", 
                                          decimal.mark = ",", 
                                          prefix = "", 
                                          suffix = " €"))
```


## Writing Files

### Excel
```{r}
bike_orderlines_wrangled_city_separated_tbl %>%
  write_xlsx("00_data/01_bike_sales/02_wrangled_data/bike_orderlines.xlsx")
```

### CSV
```{r}
bike_orderlines_wrangled_city_separated_tbl %>% 
  write_csv("00_data/01_bike_sales/02_wrangled_data/bike_orderlines.csv")
```


### RDS
```{r}
bike_orderlines_wrangled_city_separated_tbl %>% 
  write_rds("00_data/01_bike_sales/02_wrangled_data/bike_orderlines.rds")
```


## Plots
```{r}
sales_by_location_tbl %>%
  
  ggplot(aes(x = State, y = state_sales)) +
  
  geom_col(fill = "#2DC6D6") + 
  
  geom_label(aes(label = sales_formatted)) + 
  
  geom_smooth(method = "lm", se = FALSE) + 
  
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  
  labs(
    title    = "States revenue per year",
    x = " ",
    y = "Revenue"
  )
```


```{r, fig.width=14, fig.height=10}
library(ggplot2)
library(scales)

sales_by_state_year_tbl %>%
  
  ggplot(aes(x = year, y = sales, fill = State)) +
  
  geom_col() +
  
  facet_wrap(~ State) +
  
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
  labs(
    title = "Revenue by year and state",
    subtitle = "Each State has its own representation",
    fill = "States"
  )

```

# Chapter 3: Data Acquisition

## Load libraries

```{r}
library(RSQLite)
library(dplyr)
library(httr)
library(glue)
library(jsonlite)
library(keyring)
library(rvest)
library(stringr)
library(purrr)
library(xopen)
library(stringi)
library(tibble)
library(owmr)
```

## Challenge 1
### OpenWeatherMap API
```{r}
Sys.setenv(OWM_API_KEY = "c457774b4dddc07ff65800bdbf885f5f")
# get current weather data by city name
(res <- get_current("London", units = "metric") %>%
    owmr_as_tibble()) %>% names()
```
```{r}
res[, 1:6]
```

#### By City ID
```{r}
(beirut <- search_city_list("Beirut")) %>%
  as.list()
```

```{r}
get_current(beirut$id, units = "metric") %>%
  owmr_as_tibble() %>% .[, 1:6]
```
```{r}
res <- find_cities_by_geo_point(
  lat = beirut$lat,
  lon = beirut$lon,
  cnt = 5,
  units = "metric"
) %>% owmr_as_tibble()

idx <- c(names(res[1:6]), "name")
res[, idx]
```

```{r}
forecast <- get_forecast("London", units = "metric") %>%
  owmr_as_tibble()

forecast[, 1:6]
```

#### Apply functions to some columns
```{r}
funcs <- list(
  temp = round,
  wind_speed = round
)
forecast %<>% parse_columns(funcs)
```

#### Do templating
```{r}
("{{dt_txt}}h {{temp}}°C, {{wind_speed}} m/s" %$$%
    forecast) %>% head(10)
```

## Challenge 2
```{r}

url= "https://www.rosebikes.de/fahrräder/rennrad"
get_bike_data <- function(url) {
  
  html_bike_category <- read_html(url)
  
  # Get the URLs
  bike_url_tbl  <- html_bike_category %>%
    html_nodes(css = ".catalog-category-bikes__title-text") %>%
    html_text()%>%
    enframe(name = "Number", value = "Bike Name")
  
  #Descriptions
  bike_database_tbl<-bike_url_tbl %>% 
    mutate(price=html_bike_category %>%
             html_nodes(css =" .catalog-category-bikes__price-title ") %>%
             html_text())
}

bike_tableout<-get_bike_data(url)
saveRDS(bike_tableout,"Data_Acquisition_Challenge.rds")
```

### Bike Table Preview ----
```{r}
bike_tableout
```
# Chapter 4: Data wrangling

## Load libraries
```{r}
library(vroom)
library(tidyverse)
library(data.table)
library(tictoc)
```

## Import Data

### Patents
```{r}
col_types <- list(
  id = col_character(),
  date = col_date("%Y-%m-%d"),
  num_claims = col_double()
)

patent_tbl <- vroom(
  file       = "patent.tsv", 
  delim      = "\t", 
  col_types  = col_types,
  na         = c("", "NA", "NULL")
)
```

### Assignee

```{r}
col_types_assignee <- list(
  id = col_character(),
  type = col_character(),
  organization = col_character()
)

assignee_tbl <- vroom(
  file       = "assignee.tsv", 
  delim      = "\t", 
  col_types  = col_types_assignee,
  na         = c("", "NA", "NULL")
)

```

### Patents Assignee

```{r}

col_types_patent_assignee <- list(
  patent_id = col_character(),
  assignee_id = col_character()
)


patent_assignee_tbl <- vroom(
  file       = "patent_assignee.tsv", 
  delim      = "\t", 
  col_types  = col_types_patent_assignee,
  na         = c("", "NA", "NULL")
)

col_types_uspc <- list(
  patent_id = col_character(),
  mainclass_id = col_number(),
  sequence = col_number()
)


uspc_tbl <- vroom(
  file       = "uspc.tsv", 
  delim      = "\t", 
  col_types  = col_types_uspc,
  na         = c("", "NA", "NULL")
)
```

## Data Acquisition
```{r}
setDT(assignee_tbl)
setDT(patent_tbl)
setDT(patent_assignee_tbl)
setDT(uspc_tbl)

patent_tbl %>% glimpse()
assignee_tbl %>% glimpse()
patent_assignee_tbl %>% glimpse()
uspc_tbl %>% glimpse()

```

## Data Wrangling
### Question 1
```{r}

setnames(assignee_tbl, "id", "assignee_id")

combined_data <- merge(x = patent_assignee_tbl, y = assignee_tbl, by = "assignee_id")


us_patents <- combined_data %>%
  filter(type == 2)%>%
  filter(!is.na(patent_id) || !is.na(organization)) %>%
  select(-type, -assignee_id)%>% 
  group_by(organization) %>%
  count(patent_id) %>%
  select(-patent_id)%>%
  summarise(total = sum(n))%>%
  arrange(desc(total))   

us_top_10 <- us_patents %>% slice(1:10)
us_top_10

```

### Question 2
```{r}
tbl_2 <- patent_tbl %>%   
         separate(col  = date,
         into = c("year", "month", "day"),
          sep  = "-", remove = TRUE) %>%
          mutate(
              month = as.numeric(month)
            )%>%
          filter(month == 01)%>%
          select(-year, -day)

setnames(tbl_2, "id", "patent_id")
combined_data_2 <- merge(x = tbl_2, y = combined_data, by = "patent_id")

us_top10_2014_01 <- combined_data_2%>%
                    filter(type == 2)%>%
                    filter(!is.na(patent_id) || !is.na(organization)) %>%
                    select(organization, patent_id) %>%
                    group_by(organization) %>%
                    count(patent_id) %>%   
                    summarise(total_patents = sum(n))%>%
                    arrange(desc(total_patents)) %>% slice(1:10)  

us_top10_2014_01_new <- combined_data_2%>%
                        filter(type == 2 & num_claims == 1)%>%
                        filter(!is.na(patent_id) || !is.na(organization)) %>%
                        select(organization, patent_id) %>%
                        group_by(organization) %>%
                        count(patent_id) %>%   
                        summarise(total_patents = sum(n))%>%
                        arrange(desc(total_patents)) %>% slice(1:10)
us_top10_2014_01_new
```

### Question 3
```{r}
combined_data_3 <- merge(x = uspc_tbl, y = combined_data_2, by = "patent_id")



top10_worldwide_patents <- combined_data_3  %>%
                  filter(!is.na(patent_id) || !is.na(organization))%>%
                  group_by(organization) %>%
                  arrange(desc(mainclass_id)) %>% 
                  count(patent_id) %>%
                  select(-patent_id)%>%
                  summarise(total_patents_worldwide = sum(n))%>%
                  ungroup() %>%
                  arrange(desc(total_patents_worldwide)) %>% slice(1:10)  

top10_worldwide_top5 <- top10_worldwide_patents %>% slice(1:5)  
top10_worldwide_top5

```


# Chapter 5: Data Visualization
## Challenge 1
```{r}
library(tidyverse)
library(randomcoloR)
covid_data_tbl <- read_csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv")
View(covid_data_tbl)
ylab=c(0,2.5,5,7.5,10,12.5,15,17.5)
xlab=month.abb[1:12]
continents=unique(covid_data_tbl$continentExp)

#Group data
grouped1 <- covid_data_tbl %>%
  select(month, cases, continentExp) %>%
  
  group_by(month,continentExp) %>%
  summarize(
    cases = sum(cases),
  ) %>%
  ungroup() 
colnames(grouped1)=c("month","continents","cases")

#Add missing values
grouped2=data.frame(continents=rep(continents,12),stringsAsFactors = F)
grouped2=grouped2[order(grouped2$continents),]
grouped2=as.data.frame(cbind(grouped2,rep(1:12,length(continents))))
colnames(grouped2)=c("continents","month")
grouped2=merge.data.frame(grouped2,grouped1,all=T)
grouped2[is.na(grouped2)]=0

#Calculate Cumulative Cases
grouped3<- grouped2%>%
  select(month, cases, continents) %>%
  
  group_by(continents) %>%
  summarize(
    cumulativeCases = cumsum(cases),
  ) %>%
  select(cumulativeCases, continents) %>%
  ungroup()
#Add month column
grouped3=as.data.frame(cbind(grouped3,rep(1:12,length(continents))))
colnames(grouped3)=c("cumulativeCases","continents","month")
View(grouped3)

#Canvas
ggplot(data=grouped3,
       aes(x=month,y=cumulativeCases,col=continents),group=continents)+
  geom_line()+
  scale_x_continuous(name="Year 2020",breaks = 1:12,labels=xlab)+
  theme(axis.text.x = element_text(angle=45,hjust=1),axis.text.x.bottom =element_text(xlab))+
  scale_y_continuous(expand=c(0,0),name="Cumulative Cases",labels=paste0(ylab,"M"),breaks=10^6*ylab,limits = c(0,17.5e6))+
  scale_color_manual(values=randomColor(length(continents)))+
  labs(x="Year 2020",
       y="Cumulative Cases",
       title="COVID-19 confirmed cases worldwide\nAs of 11/02/2020, Europe had more cases than USA")+
  theme(plot.title = element_text(face="bold",hjust=0,vjust=2.12,size=7),plot.caption = element_text(face="bold.italic"))
```